---
title: Descriptive Analysis
author: Gary Weissman
date: December 15, 2022
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
library(data.table)
library(ggplot2)
library(lubridate)
library(skimr)
library(ggsci)
```


```{r load}
dt <- fread('fda_approved_devices.csv')
# Clean up year format issue
dt[, dofd_clean := gsub('/19$', '/2019', dt$`Date of Final Decision`)]
dt[, fin_decis_date := as.Date(dofd_clean, format = '%m/%d/%Y')]
dt[, fin_decis_moyr := floor_date(fin_decis_date, unit = 'month')]
```

```{r desc}
skim(dt)
```


```{r viz}
# Trends in final decision dates
dt[, .N, by = fin_decis_moyr] |>
  ggplot(aes(fin_decis_moyr, N)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  ggtitle('Final approval decisions each month')

# Cumulative decisions by month
dt[, .N, by = fin_decis_moyr][order(fin_decis_moyr)][, cum_decis := cumsum(N)] |>
  ggplot(aes(fin_decis_moyr, cum_decis)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  scale_x_date('Year-Month') +
  scale_y_continuous('Count') +
  ggtitle('Cumulative approval decisions by month')

# Decisions by category
dt[, .N, by = `Panel (Lead)`] |>
  ggplot(aes(reorder(`Panel (Lead)`, N), N)) +
  geom_bar(stat = 'identity', color = 'blue', fill = NA) +
  geom_text(aes(label = N, y = N + 6), color = 'black') +
  theme_bw() +
  scale_x_discrete('Panel name') +
  ggtitle('Approvals by panel') +
  coord_flip()

# Decisions by company (only the top 20)
dt[, .N, by = Company][order(-N)][1:20] |>
  ggplot(aes(reorder(Company, N), N)) +
  geom_bar(stat = 'identity', color = 'blue', fill = NA) +
  geom_text(aes(label = N, y = N + 2), color = 'black') +
  theme_bw() +
  scale_x_discrete('Company') +
  ggtitle('Company') +
  coord_flip()

```

## Hospital based

Consider the subgroup of devices related to hospital-based, critical care focused, bedside clinical care for adults.

```{r hospdev}
cc_panels <- c('Cardiovascular', 'Anesthesiology', 'General Hospital')
dt[`Panel (Lead)` %in% cc_panels]
print(paste0(nrow(dt[`Panel (Lead)` %in% cc_panels]), ' potential devices identified in these critical care categories.'))

# Look at above visualizations in this subset.
# Trends in final decision dates
dt[`Panel (Lead)` %in% cc_panels][, .N, by = fin_decis_moyr] |>
  ggplot(aes(fin_decis_moyr, N)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  ggtitle('Final approval decisions each month')

# Cumulative decisions by month
dt[`Panel (Lead)` %in% cc_panels][, .N, by = fin_decis_moyr][order(fin_decis_moyr)][, cum_decis := cumsum(N)] |>
  ggplot(aes(fin_decis_moyr, cum_decis)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  scale_x_date('Year-Month') +
  scale_y_continuous('Count') +
  ggtitle('Cumulative approval decisions by month')

# Decisions by category
dt[`Panel (Lead)` %in% cc_panels][, .N, by = `Panel (Lead)`] |>
  ggplot(aes(reorder(`Panel (Lead)`, N), N)) +
  geom_bar(stat = 'identity', color = 'blue', fill = NA) +
  geom_text(aes(label = N, y = N + 6), color = 'black') +
  theme_bw() +
  scale_x_discrete('Panel name') +
  ggtitle('Approvals by panel') +
  coord_flip()

# Decisions by company (only the top 20)
dt[`Panel (Lead)` %in% cc_panels][, .N, by = Company][order(-N)][1:20] |>
  ggplot(aes(reorder(Company, N), N)) +
  geom_bar(stat = 'identity', color = 'blue', fill = NA) +
  geom_text(aes(label = N, y = N + 2), color = 'black') +
  theme_bw() +
  scale_x_discrete('Company') +
  ggtitle('Company') +
  coord_flip()

# Number of unique products by Panel category
unique(dt[`Panel (Lead)` %in% cc_panels, .(`Primary Product Code`,
                                    `Panel (Lead)`)])[,
                                                     .N, by = .(`Panel (Lead)`)] |>
  ggplot(aes(reorder(`Panel (Lead)`, N), N)) +
  geom_bar(stat = 'identity', color = 'blue', fill = NA) +
  geom_text(aes(label = N, y = N + 1), color = 'black') +
  theme_bw() +
  scale_x_discrete('Panel name') +
  ggtitle('Approved unique devices by panel') +
  coord_flip()
```

## Now a manual review

Manually review the descriptions of these unique devices to see if they might qualify as relevant to bedside clinical decision support systems.

```{r manualreview}
#dput(unique(dt[`Panel (Lead)` %in% cc_panels, .(`Submission Number`)]))
# Here manually review each submission and indicate if it is eligible for inclusion if it relates to bedside, inpatient care in a critical care or adjacent space
# Ok this should have been done in a csv file or some database but here we are...
man_rev_list <- list(c("K210484", "No"), # implantable ECG device
                       c("DEN200038", "No"), # vital sign monitoring only
                       c("DEN200019", "No"), # optical vital sign monitoring
                       c("DEN200022", "Yes"), # provides CDS for hemodynamic monitoring
                       c("K200717", "Yes"), # ICU CDS system
                       c("K193631", "No"), # Electronic stethoscope
                       c("K201298", "No"), # atrial mapping during EP procedures
                       c("K200833", "No"), # outpatient cardiac rhythm monitor and alert system
                       c("K192732", "No"), # outpatient cardiac rhythm monitor and alert system
                       c("K200036", "Yes"), # inpatient real time ECG analysis and interpretation
                       c("K192004", "No"), # arrhythmia and heart sound/murmur analysis and detection
                       c("K183089", "No"), # hand-held spirometer
                       c("K192442", "No"), # FFR calculation from coronary angiography images
                       c("K190013", "No"), # Diabetes self-management system
                       c("K183370", "Yes"), # Pera - interface to Rothman Index for RRTs, etc
                       c("K183282", "No"), # Home monitoring of vital signs
                       c("K190925", "No"), # FFR calculation and analysis from coronary CT images
                       c("K183322", "No"), # ambulatory ECG and vitals monitoring
                       c("K183646", "Yes"), # Hypotension prediction from arterial waveforms
                       c("K181988", "No"), # electronic stethoscope heart sound analyzer
                       c("K181352", "No"), # home vital sign monitoring
                       c("K181823", "No"), # Ambulatory ECG analysis
                       c("K182344", "No"), # ECG analysis and interpretation
                       c("K183704", "No"), # Ambulatory ECG analysis
                       c("K181916", "No"), # Insulin self-management for pt with DM-II
                       c("K182456", "No"), # Watch to record single-lead ECG for outpatients
                       c("K182149", "No"), # FFR calculations from angiography images
                       c("K182790", "No"), # non acute ECG and other vitals monitoring
                       c("K180432", "Yes"), # AI-ECG is automated ECG analysis and intepretation
                       c("K173872", "No"), # Patient self-testing for atrial fibrillation
                       c("K181502", "No"), # Long-term ECG monitoring and event detection
                       c("K172959", "Yes"), # Pera health - Rothman Index interface
                       c("DEN160044", "Yes"), # Hypotension prediction monitor in OR from arterial waveform
                       c("K171056", "Yes"), # Real time EHR and waveform reviews for CDS
                       c("K172507", "No"), # Home monitoring of fluid status for HF patients
                       c("K163694", "No"), # general home monitoring of ECG
                       c("K171936", "No"), # outpatient long term ECG monitoring
                       c("K163339", "No"), # Microcirculation inspection device
                       c("K170568", "No"), # Non acute care ECG interpretation
                       c("K163665", "No"), # abdominal movement sensor to identify sleep disordered breathing; sleep but not CC
                       c("K162855", "No"), # subcutaneous implantable ECG recorder/Analysis
                       c("K160401", "No"), # cloud transmission of heart sounds via electronic stethoscope
                       c("K160016", "No"), # electronic stethoscope tool
                       c("K150102", "No"), # snoring detector
                       c("K142512", "No"), # Vital sign monitoring for ambulatory patients
                       c("K142918", "No"), # Subcutaneous insulin dose calculator
                       c("K081140", "Yes"), # vital sign monitoring and alarm suitable for inpatients
                       c("K955841", "No"), # Sleep study monitoring device
                     c("DEN200030", "No"), # Device for trans-catheter Left Atrial appendage occlusion procedures 
                     c("K152733", "No"), # Analyzing coronary arteries on CT imaging studies
                     c("K161772", "No"), # Analyzing coronary arteries on CT imaging studies
                     c("K182035", "No"), # Analyzing coronary arteries on CT imaging studies
                     c("K202527", "No"), # Analysis of long-term monitoring cardiac devices
                     c("K203329", "No"), # Analyzing coronary arteries on CT imaging studies
                     c("K210543", "No"), # Analysis of long-term monitoring cardiac devices
                     c("K210822", "No"), # Analysis of long-term. ambulatory monitoring cardiac devices
                     c("K212516", "No"), # OTC, AppleWatch arrhythmia identification
                     c("K212662", "No"), # Analysis of long-term. ambulatory monitoring cardiac devices
                     c("K213357", "No"), # Verily "study watch" to monitor outpatient abnormal cardiac rhythms with small device
                     c("K213409", "No"), # Verily  Zio "watch" to monitor outpatient abnormal cardiac rhythms with small device
                     c("K213657", "No"), # Analyzing coronary arteries on CT imaging studies
                     c("K213794", "No"), # Stethoscope murmur analysis
                     c("K213971", "No"), # Apple Watch Afib detection
                     c("K220766", "No")) # Stethoscope murmur analysis

man_rev_dt <- data.table (submission_number = sapply(man_rev_list, `[[`, 1),
                          eligible = sapply(man_rev_list, `[[`, 2))
print(paste0('After manual review there are ', sum(man_rev_dt$eligible == 'Yes'), ' approvals related to devices that could be used for acute care clinical decision support.'))
```

Now for each eligible device, record:

1. The level of evidence in the cited papers for their use (e.g. pre-post study, retrospective cohort study, use of quasi-experimental methods for causal inference, RCT, etc).
2. Whether or not the study was peer reviewed (vs pre-print)
3. Whether or not the code for the underlying algorithms are made publicly available.
4. Is there explicit assessment of effectiveness, safety (i.e harms), and equity?
5. Other features of the model related to transparency (inputs, labels, population, etc)?

```{r generatecsvtoreview}
to_review_dt <- merge(man_rev_dt, dt, all.y = TRUE,
                      by.x = 'submission_number', by.y = "Submission Number")

to_review_dt <- to_review_dt[eligible == 'Yes', .(submission_number, company = Company,
                                                  device = Device, panel = `Panel (Lead)`,
                                                  primary_product_code = `Primary Product Code`,
                                                  final_decision_dt = fin_decis_date,
                                                  study_design = '',
                                                  peer_reviewed = '',
                                                  sample_size = '',
                                                  number_of_sites = '',
                                                  code_availability = '',
                                                  safety_assessment = '',
                                                  equity_assessment = '',
                                                  other_notes = '')]

# NB. Be sure to change the name of this file once you start entering data so that you don't actually write over everything!
fwrite(to_review_dt, 'manual_review_template_empty_version.csv')
```

## Make exhibits for manuscript

```{r fig1}

fig1_dt <- merge(man_rev_dt, dt,
                      by.x = 'submission_number', by.y = "Submission Number",
                 all = TRUE)
# Identify select panels of interest
select_panels <- c('Radiology', 'Cardiovascular', 'Anesthesiology', 
                   'General Hospital', 'General and Plastic Surgery')

# and get data just for the CC CDS approvals in this study
fig1_cccds_dt <- fig1_dt[eligible == 'Yes', .N, by = .(panel = `Panel (Lead)`, fin_decis_moyr)][order(panel, fin_decis_moyr)][, cum_decis := cumsum(N), by = panel]


fig1_dt[, .N, by = .(panel = `Panel (Lead)`, fin_decis_moyr)][order(panel, fin_decis_moyr)][, cum_decis := cumsum(N), by = panel][panel %in% select_panels] |>
  ggplot(aes(fin_decis_moyr, cum_decis, color = panel)) +
  geom_line() +
  geom_point(size = 0.5) +
  theme_classic() +
  geom_point(data = fig1_cccds_dt, color = 'black', size = 0.5) +
  geom_line(data = fig1_cccds_dt, color = 'black', linetype = 'twodash') +
  scale_x_date('Date') +
  scale_y_continuous('Count') +
  annotate('text', as.Date('2006-01-01'), 100, 
           label = 'Critical care \nclinical decision support') +
  annotate("segment", x = as.Date('2003-01-15'), 
           xend = as.Date('2008-12-31'), 
           y = 80, 
           yend = 80,
           colour = "black",
           linetype = 'twodash') + 
  annotate('point', x = as.Date('2006-01-01'), y = 80, color = 'black') + 
  #ggtitle('Cumulative FDA approvals for devices using AI or ML technology') +
  scale_color_nejm(name = 'FDA review panel') +
  theme(legend.position = 'bottom')

ggsave('exhibits/fig1_temporal_trends.png')

```